---
title: "The size premium in private equity"
description: |
  A short description of the post.
author:
  - name: Christoph Jäckel
    url: {}
date: 02-14-2021
output:
  distill::distill_article:
    self_contained: false
    code_folding: true
categories:
  - Private Equity
draft: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(data.table)
library(ggplot2)
library(knitr)
library(kableExtra)
options(scipen=999) # https://stackoverflow.com/questions/5352099/how-to-disable-scientific-notation
```

Starting with the size premium in public equity, refer to your value bridge article and start with the premise that many smaller buyout managers have a simple value creation story: increase the EBITDA and take a multiple arbitrage, kinda implying that just for a larger business, people pay more. Go back then to financial theory 101, value additivity: http://www.its.caltech.edu/~rosentha/courses/BEM103/Readings/BOCh03.pdf

The price of a basket of goods (apples, pears,...) equals the sum of the prices of the individual goods.

So just by bringing to businesses together, you are not magically increasing its value. Argue than how value can created.

Then show examples with a value bridge...


```{r Value_Bridge_function, include=FALSE}
value_creation_LBO <- function(.startEquity, .endEquity, .startDebt, .endDebt, .startRev, .endRev, .startEBITDA, .endEBITDA,
                               .interimCC, .interimDist, .interimIntRate, .holdingPeriod) {

  ### Calculations of additional inputs
  startEV <- .startEquity + .startDebt
  endEV   <- .endEquity   + .endDebt
  startEV_EBITDA_mult <- startEV/.startEBITDA
  endEV_EBITDA_mult   <- endEV/.endEBITDA
  costDebt <- (1+.interimIntRate)^.holdingPeriod - 1
  invCapital <- .startEquity + .interimCC
  deltaEquity <- .endEquity - .startEquity
  gain <- deltaEquity + .interimDist - .interimCC
  avgDebtEquity <- (.startDebt/.startEquity + .endDebt/.endEquity)/2 #TODO: Understand why you need the average here
  startMargin <- .startEBITDA/.startRev
  endMargin   <- .endEBITDA/.endRev

  ### Calculations of output
  TM_lev                  <- gain/invCapital
  TM_unlev                <- (TM_lev + costDebt * avgDebtEquity)/(1 + avgDebtEquity)
  lev_effect              <- TM_lev - TM_unlev
  mult_effect             <- .startEBITDA * (endEV_EBITDA_mult - startEV_EBITDA_mult)
  mult_EBITDA_comb_effect <- (.endEBITDA - .startEBITDA) * (endEV_EBITDA_mult - startEV_EBITDA_mult)
  FCF_effect              <- -(.endDebt - .startDebt) + .interimDist - .interimCC
  EBITDA_effect           <- (.endEBITDA - .startEBITDA) * startEV_EBITDA_mult
  Rev_effect              <- (.endRev - .startRev) * startMargin * startEV_EBITDA_mult
  margin_effect           <- (endMargin - startMargin) * .startRev  * startEV_EBITDA_mult
  SM_comb_effect          <- (.endRev - .startRev) * (endMargin - startMargin) * startEV_EBITDA_mult

  vec_results <-c(TM_lev                  = TM_lev,
                  TM_unlev                = TM_unlev,
                  gain                    = gain,
                  lev_effect              = lev_effect,
                  mult_effect             = mult_effect,
                  mult_EBITDA_comb_effect = mult_EBITDA_comb_effect,
                  EBITDA_effect           = EBITDA_effect,
                  FCF_effect              = FCF_effect,
                  Rev_effect              = Rev_effect,
                  margin_effect           = margin_effect,
                  SM_comb_effect          = SM_comb_effect,
                  startEV_EBITDA_mult     = startEV_EBITDA_mult,
                  endEV_EBITDA_mult       = endEV_EBITDA_mult,
                  startEV                 = startEV,
                  endEV                   = endEV,
                  totalInv                = .startEquity + .interimCC,
                  TM_mult                 = mult_effect/abs(gain) * abs(TM_unlev),
                  TM_mult_EBITDA_comb     = mult_EBITDA_comb_effect/abs(gain) * abs(TM_unlev),
                  TM_EBITDA               = EBITDA_effect/abs(gain) * abs(TM_unlev),
                  TM_FCF                  = FCF_effect/abs(gain) * abs(TM_unlev),
                  TM_Rev                  = Rev_effect/abs(gain) * abs(TM_unlev),
                  TM_margin               = margin_effect/abs(gain) * abs(TM_unlev),
                  TM_SM_comb              = SM_comb_effect/abs(gain) * abs(TM_unlev))
  
    print_table_inputs <- data.table(
     Item  = c("Revenue", "EBITDA", "Equity", "Net Debt", "EV", "EV/EBITDA", "CC", "Dist",
               "Holding period", "Interest rate p.a.",  "Avg debt/equity ratio", "Cost of debt"),
     Entry = c(.startRev, .startEBITDA, .startEquity, .startDebt, startEV, startEV_EBITDA_mult,
               rep(NA, 6)),
     Interim = c(rep(NA, 6), .interimCC, .interimDist, 
                 .holdingPeriod, .interimIntRate, avgDebtEquity, costDebt),
     Exit    = c(.endRev, .endEBITDA, .endEquity, .endDebt, endEV, endEV_EBITDA_mult, rep(NA, 6))
  )
  
  print_table_outputs <- data.table(
     Item  = c("TM (levered)", 
               "Gain / TM (unlevered)",
               "FCF",
               "Comb. multiple / EBITDA",
               "Multiple",
               "EBITDA",
               "Comb. revenue/margin",
               "Revenue",
               "Margin"),
     Absolute = c(NA, vec_results[c(3,8,6,5,7,11,9,10)]),
     TM       = c(vec_results[c(1,2,20,18,17,19,23,21,22)])
  )
  
  
  ### Output
  return(list(vec_results         = vec_results,
              print_table_inputs  = print_table_inputs,
              print_table_outputs = print_table_outputs))

}
```


```{r Plotting_value_bridge, include=FALSE}
plot_value_bridge <- function(.vec_value_creation_LBO, .fillColors = c("#FFC300", "#581845")) {

  #Calculate inbetween multiples
  mult_OpImpr <- .vec_value_creation_LBO["TM_unlev"] - .vec_value_creation_LBO["TM_mult"] - .vec_value_creation_LBO["TM_mult_EBITDA_comb"]

  strDesc <- c("Combination S/M effect", "Margin effect", "Sales effect", "EBITDA effect", "FCF effect", "Operating improvements",
               "Combination effect", "Multiple effect", "TM (unlev)", "Leverage effect", "TM (lev)")

  plotDF <- data.frame(Desc  = factor(strDesc, levels = strDesc),
                       Type  = c("Bridge", "Bridge", "Bridge", "Full", "Bridge", "Full", "Bridge", "Bridge", "Full", "Bridge", "Full"),
                       Start = c(0,
                                 .vec_value_creation_LBO["TM_SM_comb"], 
                                 .vec_value_creation_LBO["TM_SM_comb"] + .vec_value_creation_LBO["TM_margin"], 
                                 0,
                                 .vec_value_creation_LBO["TM_EBITDA"], 
                                 0,
                                 mult_OpImpr,
                                 mult_OpImpr + .vec_value_creation_LBO["TM_mult_EBITDA_comb"], 
                                 0,
                                 .vec_value_creation_LBO[["TM_unlev"]],
                                 0),
                       End   = c(.vec_value_creation_LBO["TM_SM_comb"],
                                 .vec_value_creation_LBO["TM_SM_comb"] + .vec_value_creation_LBO["TM_margin"],
                                 .vec_value_creation_LBO["TM_EBITDA"],
                                 .vec_value_creation_LBO["TM_EBITDA"],
                                 mult_OpImpr,
                                 mult_OpImpr,
                                 mult_OpImpr + .vec_value_creation_LBO["TM_mult_EBITDA_comb"],
                                 .vec_value_creation_LBO[["TM_unlev"]],
                                 .vec_value_creation_LBO[["TM_unlev"]],
                                 .vec_value_creation_LBO[["TM_lev"]],
                                 .vec_value_creation_LBO[["TM_lev"]]))
  plotDF$Amount <- plotDF$End - plotDF$Start
  plotDF$ID <- 1:nrow(plotDF)

  #https://stackoverflow.com/questions/50688764/r-ggplot2-ignoring-unknown-aesthetics-with-geom-rect
  suppressWarnings(ggplot(plotDF, aes(x=Desc, fill = Type)) +
                     scale_fill_manual(values = .fillColors) +
                     geom_rect(aes(x=Desc, xmin = ID - 0.45, xmax = ID + 0.45, ymin = Start, ymax = End)) +
                     xlab("") +
                     theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust = 1)))

}

```


## The starting point: no value creation

Start with example, left pocket, right pocket. Doesn't create value.

```{r Start_left_pocket_right_pocket, echo=FALSE}
interest <- 0.00 #As you simply do an analysis before/after acquisition
### Start with company
startEV     <- 100
startEquity <- 50
startDebt   <- startEV - startEquity
startEBITDA <- 10
startRev    <- 100
startMargin <- startEBITDA/startRev
EV_EBITDA_multiple <- startEV/startEBITDA
### Acquire business with the same characteristics for the same multiple
#   Assume it's fully financed with debt
addEV <- 100
addEBITDA <- addEV/EV_EBITDA_multiple
addRev <- addEBITDA / startMargin


list_results <- value_creation_LBO(
                   .startEquity         = startEquity,
                   #As additional EV is fully financed by debt, I add and subtract it; 0.01 needed otherwise NA
                   .endEquity           = startEV + addEV - startDebt - addEV + 0.01, 
                   .startDebt           = startDebt,
                   .endDebt             = startDebt + addEV,
                   .startRev            = startRev,
                   .endRev              = startRev + addRev,
                   .startEBITDA         = startEBITDA,
                   .endEBITDA           = startEBITDA + addEBITDA,
                   .interimCC           = 0,
                   .interimDist         = 0,
                   .interimIntRate      = interest,
                   .holdingPeriod       = 0)
kbl(list_results$print_table_inputs,
    caption="Inputs of add-on example") %>%
  kable_classic(full_width = FALSE)
kbl(list_results$print_table_outputs,
    caption="Results of value creation analysis of add-on example") %>%
  kable_classic(full_width = FALSE)
# Doesn't make sense here, as no value created
plot_value_bridge(list_results$vec_results)
```

## Synergies 

Now bring synergies into the game. 


```{r Synergies, echo=FALSE}
### Assume margin improvement of 5%, same multiple
margin_impr <- 0.05
endEBITDA <- startEBITDA + addEBITDA + margin_impr*(startRev + addRev)
endEV <- endEBITDA * EV_EBITDA_multiple
endEquity <- endEV - (startDebt + addEV)

list_results <- value_creation_LBO(
                   .startEquity         = startEquity,
                   #As additional EV is fully financed by debt, I add and subtract it; 0.01 needed otherwise NA
                   .endEquity           = endEquity, 
                   .startDebt           = startDebt,
                   .endDebt             = startDebt + addEV,
                   .startRev            = startRev,
                   .endRev              = startRev + addRev,
                   .startEBITDA         = startEBITDA,
                   .endEBITDA           = endEBITDA,
                   .interimCC           = 0,
                   .interimDist         = 0,
                   .interimIntRate      = interest,
                   .holdingPeriod       = 0)
kbl(list_results$print_table_inputs,
    caption="Inputs of add-on example with synergies") %>%
  kable_classic(full_width = FALSE)
kbl(list_results$print_table_outputs,
    caption="Results of value creation analysis of add-on example with synergies") %>%
  kable_classic(full_width = FALSE)
# Doesn't make sense here, as no value created
plot_value_bridge(list_results$vec_results)
```

## Roll-ups

Describe [Roll ups](https://en.wikipedia.org/wiki/Rollup) as a special form, in which multiple arbitrage is reasonable to assume, simply because you are providing liquidity to an otherwise illiquid market (e.g., very difficult for a dentist to sell). With everything in life though, it's supply and demand and if there are too many players active in th market, it also becomes more and more difficult to profit from this illiquidity premium.

Also, it is not a trivial question, and one that does by far not get as much attention as it should, to have a reasonable estimate on how large this multiple expansion should be. 2x, 4x, 6x? Driven by what factors?

Hence, it is important again to link the analysis back to free cash flows (link to your [blog post](https://www.christoph-jaeckel.com/posts/2021-01-17-evebitda-valuation-multiple-the-theory/)).